---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include= TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
```

## Loading and preprocessing the data

```{r loading, include= TRUE}

download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",
              destfile = "repdata_data_activity.zip")

df<- unzip(zipfile = "repdata_data_activity.zip")

df<- read_csv(df)


```

## What is mean total number of steps taken per day?

### Histogram of the total number of steps taken each day

```{r hist}

df %>%
  group_by(date) %>% 
  summarise(n_steps= sum(steps, na.rm= F)) %>% 
  ggplot(mapping = aes(x= n_steps))+
  geom_histogram(binwidth = 5000, color= "black", fill= "grey80")+
  xlab("Number of steps")+
  ylab("Frequency")+
  ggtitle("Total Steps per Day")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

### Mean and median number of steps taken each day

```{r mean_median}

mean<- df %>%
  group_by(date) %>% 
  summarise(n_steps= sum(steps, na.rm= F)) %>% 
  ungroup() %>% 
  summarise(avg_steps = mean(n_steps, na.rm = T)) %>% 
  pull()

mean

median<- df %>%
  group_by(date) %>% 
  summarise(n_steps= sum(steps, na.rm= F)) %>% 
  ungroup() %>% 
  summarise(med_steps = median(n_steps, na.rm = T)) %>% 
  pull()

median
```


## What is the average daily activity pattern?

### Time series plot of the average number of steps taken

```{r time_series}

df %>%
  group_by(interval) %>% 
  summarise(n_steps= mean(steps, na.rm= T))  %>% 
  ggplot(mapping = aes(x= interval, y= n_steps))+
  geom_line()+
  xlab("Interval")+
  ylab("Number of Steps")+
  ggtitle("Average Steps per Day by Interval")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
  
```

### The 5-minute interval that, on average, contains the maximum number of steps

```{r max_int}

max_int<- df %>%
  group_by(interval) %>% 
  summarise(n_steps= mean(steps, na.rm= T)) %>% 
  arrange(desc(n_steps)) %>% 
  head(1)

max_int[, 1]
  
```

## Imputing missing values

```{r missing_data}

sum(is.na(df)) # Calculate and report the total number of missing values in the dataset

fill_df <- df %>%
  group_by(interval) %>%  
  mutate(steps = ifelse(is.na(steps), 
                        mean(steps, na.rm = TRUE), # I used  the mean for that 5-minute interval
                        steps)) %>% 
  ungroup()

head(fill_df)

```


### Histogram of the total number of steps taken each day after missing values are imputed

```{r fill_hist}

fill_df %>%
  group_by(date) %>% 
  summarise(n_steps= sum(steps, na.rm= F)) %>% 
  ggplot(mapping = aes(x= n_steps))+
  geom_histogram(binwidth = 5000, color= "black", fill= "grey80")+
  xlab("Number of steps")+
  ylab("Frequency")+
  ggtitle("Total Steps per Day")+
  labs(caption = "*The average for the corresponding interval was used to fill the null observations")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


## Are there differences in activity patterns between weekdays and weekends?

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r weekends}

df %>% 
  mutate(w_day= wday(date, label = T, abbr = F),
         w_day= case_when(w_day == "sábado" ~ "Weekend",
                          w_day == "domingo" ~ "Weekend",
                          T ~ "Weekday")) %>%
  summarise(.by = c(w_day, interval),
            steps= mean(steps, na.rm= T)) %>% 
  ggplot(mapping = aes(x= interval, y= steps))+
  geom_line()+
  xlab("Interval")+
  ylab("Number of steps")+
  ggtitle(label = "Total Steps per Day by Interval",
          subtitle = "Weekdays vs Weekends")+
  facet_wrap(~w_day, nrow= 2)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5))
```